---
title: "R Notebook"
output: 
  html_notebook: 
    highlight: espresso
    number_sections: yes
    theme: journal
---

```{r libraries}
library(cluster)
library(mclust)
library(lattice)
library(tidyverse)
library(maptree)
library(mice)
library(magrittr)
library(mclust)
library(mice)
library(nFactors)
library(poLCA)
library(Rtsne)
library(corrplot)
```

# Utility Functions
```{r}
utility_histogramplotter <- function(x) {
  lattice::histogram( ~ x)
}
seg.summ <- function(data, groups) {
  aggregate(data, list(groups), function(x) mean(as.numeric(as.character(x))))
}
minmax <- function(x)
{
    return((x- min(x)) /(max(x)-min(x)))
}
plot_bi_plots <- function(df, y,...) {
  old.par = par()
  par(mar=c(5,4,5,2))
  on.exit(expr = 'par = old.par')
  prcomp(df) %>% biplot(col = c('gray', 'red'), cex = .8, main = y, ...)
}
adjust_q13 <- function(df){
  df %<>% 
    mutate(q13_visitfreq_social= (q13r1+q13r2+q13r3 +q13r11)/4,
           q13_visitfreq_music = (q13r4+q13r7+q13r8 +q13r9)/4,
           q13_visitfreq_video = (q13r5+q13r6+q13r10+q13r12)/4)
  df[,!str_detect(names(df),'q13r')]
}
adjust_q24 <- function(df){
  df %<>% 
    mutate(
      q24_tech_posatt = (q24r1+q24r2+q24r3+q24r5+q24r6)/5,
      q24_tech_enter = (q24r7+q24r8)/2,
      q24_tech_comm = (q24r10+q24r11+q24r12)/3,
      q24_tech_negatv = (q24r4+q24r9)/2
    )
  df[,!str_detect(names(df),'q24r')]
}
adjust_q25 <- function(df){
  df %<>% 
    mutate(
      q25_prsnlty_leader = (q25r1+q25r2+q25r3+q25r4)/4,
      q25_prsnlty_risk   = (q25r5+q25r7+q25r8)/3,
      q25_prsnlty_drive  = (q25r9+q25r10+q25r11+q25r12)/4,
      q25_prsnlty_follower = q25r6
    )
  df[,!str_detect(names(df),'q25r')]
}
adjust_q26 <- function(df){
  df %<>% 
    mutate(
      q26_shopsavvy_bargain =  (q26r3+q26r5)/2,
      q26_shopsavvy_brands = (q26r18+q26r7+q26r13+q26r14+q26r15)/5,
      q26_shopsavvy_earn2spend = (q26r13+q26r16+q26r4)/3,
      q26_shopsavvy_applover = (q26r17+q26r12+q26r10+q26r8+q26r6+q26r9)/6,
      q26_shopsavvy_children = q26r11,
    )
  df[,!str_detect(names(df),'q26r')]
}
adjust_q2 <- function(df){
  df %<>%
    mutate(
      q2_apple = ifelse(q2r1==1|q2r2==1,1,0),
      q2_andriod = q2r3,
      q2_windows = q2r6,
      q2_tablet = q2r8,
      q2_other = ifelse(q2r4==1|q2r5==1|q2r7==1|q2r9==1,1,0)
    )
  df[,!str_detect(names(df),'q2r')]
}
adjust_q4 <- function(df){
    df %<>%
    mutate(
      q4_use_music_apps =q4r1,
      q4_use_tv_apps =ifelse(q4r2==1|q4r3==1|q4r4==1,1,0),
      q4_use_game_apps =q4r5,
      q4_use_social_apps =q4r6,
      q4_use_news_apps =ifelse(q4r7==1|q4r9==1,1,0),
      q4_use_shop_apps =q4r8,
      q4_use_none_apps =q4r11
    )
  df[,!str_detect(names(df),'q4r')]
}
adjust_race <- function(df){
  df %>% 
    mutate(
      q54_white  = ifelse(q54==1,1,0),
      q54_black  = ifelse(q54==2,1,0),
      q54_asian  = ifelse(q54==3,1,0),
      q54_hawai  = ifelse(q54==4,1,0),
      q54_native =  ifelse(q54==5,1,0),
      q54_other  = ifelse(q54==6,1,0),
      q55_latino =  ifelse(q55==1,1,0)
    ) %>% 
    dplyr::select(-q54,-q55)
}
adjust_gender <- function(df){
  df %>% 
    mutate(q57 = ifelse(q57==1,1,0))
}
adjust_marital <- function(df){
  df %>% 
    mutate(q49 = ifelse(q49==1,1,0))
}
adjust_q11 <- function(df){
  #Q11 has artificial ordinality between #of apps increasing, and response==5 as "Dont know"
  #This function resolves this problem
  df %>% 
    mutate(q11 = ifelse(q11==5 | q11==6, 0, q11))
}
adjust_children <- function(df){
  df %>% 
    dplyr::select(-q50r2_inftod,-q50r3_6_12,-q50r4_13_17,-q50r5_adult)
}
adjust_income <- function(df){
  df %>% 
    dplyr::mutate(q56 = case_when(
      q56 <= 4 ~ 1,
      q56 >= 5 & q56 <=8 ~ 2,
      q56 >= 9 & q56 <=11 ~ 3,
      q56 >= 12 & q56 <= 13 ~ 4,
      q56 >= 14 ~ 5
    ))
}
adjust_age <- function(df){
  df %>% 
    dplyr::mutate(
      q1 = case_when(
        q1 <= 2 ~ 1,
        q1 >= 3 & q1 <= 5 ~ 2,
        q1 >= 6 & q1 <= 8 ~ 3,
        q1 >= 9 & q1 <= 11 ~ 4,
        q1 >= 12 ~ 5,
      )
    )
}
adjust_names <- function(df){
  df %>% 
    dplyr::rename(
      q1_age=q1,
      q11_appnum = q11,
      q12_freeapppc = q12,
      q48_edu = q48,
      q49_marital = q49,
      q56_income = q56,
      q57_mf = q57,
      q50r1_nochild = q50r1,
      q50r2_inftod = q50r2, 
      q50r3_6_12 = q50r3,
      q50r4_13_17 = q50r4,
      q50r5_adult = q50r5
    )
}
```


# Data Prep
```{r}
load('../data/apphappyData.RData')
df_labs <- tbl_df(apphappy.3.labs.frame)
df_nums <- tbl_df(apphappy.3.num.frame)
rm(apphappy.3.num.frame); rm(apphappy.3.labs.frame)
df_labs$q57 <- as.factor(df_labs$q57)
df_labs$caseID <- NULL
df_nums$caseID <- NULL
df_labs$q2r10 <- NULL
df_nums$q2r10 <- NULL
df_labs$q5r1 <- NULL
df_nums$q5r1 <- NULL
```

<!-- # EDA -->

<!-- ## Summaries -->

<!-- ```{r} -->
<!-- glimpse(df_labs) -->
<!-- skimr::skim(df_labs) -->
<!-- ``` -->

<!-- ## Age -->

<!-- ```{r} -->
<!-- histogram(~q1, df_labs) -->
<!-- ``` -->

<!-- ## How many apps? -->

<!-- ```{r} -->
<!-- df_labs %>% dplyr::select(contains('q11')) %>% lattice::histogram(~q11,.) -->
<!-- ``` -->

<!-- ## PC free? -->

<!-- ```{r} -->
<!-- df_labs %>% dplyr::select(contains('q12')) %>% lattice::histogram(~q12,.) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- xtabs(formula = ~q11+q12,data = df_labs) -->
<!-- ``` -->


<!-- ## Website visits -->

<!-- ```{r} -->
<!-- df_labs %>% select(contains('q13r')) %>% map(~utility_histogramplotter(.x)) -->
<!-- ``` -->
<!-- ## Q24 -->

<!-- ```{r} -->
<!-- df_labs %>% select(contains('q24r')) %>% map(~utility_histogramplotter(.x)) -->
<!-- df_labs %>% select(contains('q24r')) %>% map_df(~as.numeric(.x)) %>% cor() %>% corrplot::corrplot(order = 'hclust') -->
<!-- ``` -->
<!-- ## Q25 -->

<!-- ```{r} -->
<!-- df_labs %>% select(contains('q25r')) %>% map(~utility_histogramplotter(.x)) -->
<!-- df_labs %>% select(contains('q25r')) %>% map_df(~as.numeric(.x)) %>% cor() %>% corrplot::corrplot(order = 'hclust') -->
<!-- ``` -->

<!-- ## Q26 -->

<!-- ```{r} -->
<!-- df_labs %>% select(contains('q26r')) %>% map(~utility_histogramplotter(.x)) -->
<!-- df_labs %>% select(contains('q26r')) %>% map_df(~as.numeric(.x)) %>% cor() %>% corrplot::corrplot(order = 'hclust') -->
<!-- ``` -->

<!-- ## Q5 -->

<!-- ```{r} -->
<!-- df_labs %>% dplyr::select(contains('q5r1')) %>% lattice::histogram(~q5r1,.,scales=list(x=list(rot=30))) -->
<!-- ``` -->

<!-- ## Q48 -->

<!-- ```{r} -->
<!-- df_labs %>% dplyr::select(contains('q48')) %>% lattice::histogram(~q48,.,scales=list(x=list(rot=30))) -->
<!-- ``` -->

<!-- ## Q49 -->

<!-- ```{r} -->
<!-- df_labs %>% dplyr::select(contains('q49')) %>% lattice::histogram(~q49,.,scales=list(x=list(rot=30))) -->
<!-- ``` -->

<!-- ## Q50 -->

<!-- ```{r} -->
<!-- df_labs %>% select(contains('q50r')) %>% map(~utility_histogramplotter(.x)) -->
<!-- ``` -->

<!-- ## Q56 -->

<!-- ```{r} -->
<!-- df_labs %>% dplyr::select(contains('q56')) %>% lattice::histogram(~q56,.,scales=list(x=list(rot=30))) -->
<!-- ``` -->




# Remove missing values

## Handle the case of the missing apps?
```{r}
df_nums$q12[is.na(df_nums$q12)] <- 0
df_nums %>% xtabs(~q12+q11,.,addNA = T)
df_nums %>% xtabs(~q4r10+q11,.,addNA = T)
df_nums %>% xtabs(~q4r11+q11,.,addNA = T)
# RULE (A): IF q4r11=TRUE, it means you have no apps. So that means q11 has to equal 6, i.e. NONE. Data shows this is violated. Correcting using this rule.
df_nums$q11[df_nums$q4r11==TRUE] <- 6
df_nums %>% xtabs(~q4r11+q11,.,addNA = T)
```

## Small changes

Since q11 can be ordinal, "none" should equal 0 instead of 6 to preserve ordinality
```{r}
# RULE (B): Set q11=6 to q11=0
df_nums$q11[df_nums$q11==6] <- 0
df_nums %>% xtabs(~q11,., addNA = T)
```

'Dont know' doesn't add value. Perhaps these can be set to `NA` and then imputed using `mice`.

```{r}
# RULE (C): Set q11 Dont know to NA
df_nums$q11[df_nums$q11==0] <- NA
df_nums %>% xtabs(~q11,., addNA = T)
```

```{r}
# RULE (D): All NA values for q12 set to 6, since I'm approximating that if you don't have any apps, might as well could as free?
df_nums$q12[is.na(df_nums$q12)] <- 6
df_nums %>% xtabs(~q12,., addNA = T)
```

```{r}
map_int(df_nums,~sum(is.na(.x))) %>% sort() %>% tail(3) %>% barchart(main='Missing values')
```


## Imputing using `mice`
```{r}
map_df(df_nums,~as.factor(.x)) %>% 
  mice::mice(printFlag = T, m = 5, method = 'rf') -> miceFit
df_nums <- tbl_df(mice::complete(miceFit))
```

# Bi Plots

```{r}
# questions_to_consolidate <- c('q13','q24','q25','q26')
# plot_bi_plots(df_nums %>% dplyr::select(contains(questions_to_consolidate[1])),questions_to_consolidate[1])
# plot_bi_plots(df_nums %>% dplyr::select(contains(questions_to_consolidate[2])),questions_to_consolidate[2], xlim=c(0,.05), ylim=c(-0.01,0.002))
# plot_bi_plots(df_nums %>% dplyr::select(contains(questions_to_consolidate[3])),questions_to_consolidate[3])
# plot_bi_plots(df_nums %>% dplyr::select(contains(questions_to_consolidate[4])),questions_to_consolidate[4], xlim=c(0,0.05),ylim=c(-0.011,0.005))
```

# Data Prep

## Sub grouping

```{r}
df_nums <- map_df(df_nums,~as.numeric(as.character(.x)))
df_nums_adj <- df_nums %>%  
  adjust_q13() %>% 
  adjust_q24() %>% 
  adjust_q25() %>% 
  adjust_q26() %>% 
  adjust_q2() %>% 
  adjust_q4() %>% 
  adjust_race() %>% 
  adjust_gender() %>% 
  adjust_marital() %>% 
  adjust_income() %>%
  adjust_age() %>% 
  adjust_q11() %>% 
  adjust_names() %>% 
  adjust_children()

glimpse(df_nums_adj)
```

## Selection of basis variables

```{r}
set1 <- c(
  'q1_age',
  # 'q11_appnum',
  # 'q12_freeapppc',
  'q48_edu',
  'q49_marital',
  'q50r1_nochild',
  'q56_income',
  # 'q57_mf',
  # 'q13_visitfreq_social',
  # 'q13_visitfreq_music',
  # 'q13_visitfreq_video',
  'q24_tech_posatt',
  'q24_tech_enter',
  'q24_tech_comm',
  'q24_tech_negatv',
  # 'q25_prsnlty_leader',
  # 'q25_prsnlty_risk',
  # 'q25_prsnlty_drive',
  # 'q25_prsnlty_follower',
  'q26_shopsavvy_bargain',
  'q26_shopsavvy_brands',
  'q26_shopsavvy_earn2spend',
  'q26_shopsavvy_applover',
  'q26_shopsavvy_children',
  'q2_apple',
  'q2_andriod',
  'q2_windows',
  # 'q2_tablet',
  # 'q2_other',
  # 'q4_use_music_apps',
  # 'q4_use_tv_apps',
  # 'q4_use_game_apps',
  # 'q4_use_social_apps',
  # 'q4_use_news_apps',
  # 'q4_use_shop_apps',
  # 'q4_use_none_apps',
  'q54_white',
  'q54_black',
  'q54_asian',
  'q54_hawai',
  'q54_native',
  # 'q54_other',
  'q55_latino'
)
```

```{r}
df_nums_adj_backup <- df_nums_adj
# df_nums_adj <- df_nums_adj_backup
```


```{r}
df_nums_adj <- df_nums_adj %>% dplyr::select(one_of(set1))
```


# Scaling
```{r}
scaling_wanted <- T
minmax_wanted <- F
if (scaling_wanted) {
  df_nums_adjscaled <- scale(df_nums_adj)
  centers <- attributes(df_nums_adjscaled)[[3]]
  spreads <- attributes(df_nums_adjscaled)[[4]]
  df_nums_adjscaled <- tbl_df(df_nums_adjscaled)
}
if(minmax_wanted){
  df_nums_adjscaled <- map_df(df_nums_adj, ~minmax(.x))
}
if(!scaling_wanted & !minmax_wanted){
  df_nums_adjscaled <- df_nums_adj
}
head(df_nums_adjscaled)
```

# Correlation Plot
```{r}
df_nums_adjscaled %>% cor() %>% corrplot(method = 'shade',tl.col = 'black',tl.cex = .9, order = 'hclust', hclust.method = 'ward.D2')
```

# Dissimilarity Calculations

```{r}
# ordered=c(1:4,11,13:28)
# symm_bin=c(6:10,12,29:47)
# df_nums_adjscaled[ordered] <- map_df(df_nums_adjscaled[ordered],~as.ordered(.x))
# df_nums_adjscaled[symm_bin] <- map_df(df_nums_adjscaled[symm_bin],~as.factor(.x))
distMat <- daisy(df_nums_adjscaled)
```


# Simple clustering

## H clust

### How cophenetic changes with clustering method?
```{r}
methods <- c('complete','average','ward.D','ward.D2','median','mcquitty','centroid')
method_vs_cop <- map_dbl(methods,~cor(cophenetic(hclust(d = distMat, method = .x)), distMat))
names(method_vs_cop) <- methods
barchart(sort(method_vs_cop))
```

```{r}
k = 4

hclustFit <- hclust(d = distMat, method = 'ward.D2')
plot(hclustFit, labels = F)
rect.hclust(hclustFit, k=k, border="red")
```


```{r}
hclust_segments <- cutree(hclustFit, k = k)
table(hclust_segments)
clusplot(df_nums_adjscaled, hclust_segments, color=TRUE, shade=TRUE, labels=4, lines=0, main="HClust plot")
```

```{r}
seg.summ(df_nums_adj, hclust_segments) -> centroids
centroids
```

```{r}
cutFit <- cutree(hclustFit, k)
plot(silhouette(cutFit,distMat))
```

## K-means

### How many clusters to use?

```{r}
wssplot <- function(numsub, nc=15, seed=222) {
  wss <- (nrow(numsub)-1)*sum(apply(numsub,2,var))
  for (i in 2:nc) {
    set.seed(seed)
    wss[i] <- sum(kmeans(numsub, centers=i, iter.max = 1e4)$withinss)}
  plot(1:nc, wss, type="b", xlab="Number of Clusters",
       ylab="Within groups sum of squares")}
wssplot(df_nums_adjscaled)
```

### Solving the k-means model

```{r}
# Using cluster centers from hclust:
seg.summ(df_nums_adjscaled, hclust_segments) -> kmeans_centroids
clusterresults <- kmeans(x = df_nums_adjscaled, centers = kmeans_centroids[,-1])
rsquare <- clusterresults$betweenss/clusterresults$totss
cat('\nWithin SS:',clusterresults$withinss,' Sizes:\n',clusterresults$size,'\nrsq:', rsquare)
clusplot(df_nums_adjscaled, clusterresults$cluster, color=TRUE, shade=TRUE, labels=4, lines=0, main="K-means cluster plot")

# Using number of clusters + random starts
k_kmeans = 5
clusterresults <- kmeans(x = df_nums_adjscaled, centers = k_kmeans)
rsquare <- clusterresults$betweenss/clusterresults$totss
cat('\nWithin SS:',clusterresults$withinss,' Sizes:\n',clusterresults$size,'\nrsq:', rsquare)
clusplot(df_nums_adjscaled, clusterresults$cluster, color=TRUE, shade=TRUE, labels=4, lines=0, main="K-means cluster plot")
```

```{r}
seg.summ(df_nums_adj, clusterresults$cluster) -> segsummary_results
segsummary_results
```

## PAM

```{r}
pamFits <- tibble(k_pam = 2:10)
pamFits$pamFits <- map(pamFits$k_pam, ~pam(df_nums_adjscaled, k = .x))
pamFits$sil_avg_width <- map_dbl(pamFits$pamFits,~.x$silinfo$avg.width)
pamFits
```

```{r}
pamFits %>% 
  xyplot(sil_avg_width~as.factor(k_pam),.,type='b')
```

```{r}
plot(pamFits$pamFits[[4]])
```

```{r}
seg.summ(df_nums_adj, pamFits$pamFits[[4]]$clustering)
```

# Model based

```{r}
mclustFits <- tibble(mclust_clusters=2:8)
mclustFits$mclustFits <- map(mclustFits$mclust_clusters, ~Mclust(df_nums_adjscaled, G = .x))
mclustFits$bic <- map_dbl(mclustFits$mclustFits, ~.x[['bic']])
mclustFits$loglik <- map_dbl(mclustFits$mclustFits, ~.x[['loglik']])
mclustFits
```

```{r}
mclustFits %>% 
  xyplot(bic+loglik~as.factor(mclust_clusters),., auto.key = T, type = 'b')
```

```{r}
summary(mclustFits[[2,'mclustFits']])
```

```{r}
mclust_clusters <- mclustFits[[2,'mclustFits']]$classification
clusplot(df_nums_adj, mclust_clusters, color=TRUE, shade=TRUE, labels=4, lines=0, main="M Clust Plot, k = 4")
```

```{r}
seg.summ(df_nums_adj, mclust_clusters)
```
```{r}
mclust_distMat <- daisy(df_nums_adj)
plot(silhouette(mclust_clusters,mclust_distMat))
```

<!-- ## tsne -->

<!-- ```{r} -->
<!-- tsneFit <- Rtsne::Rtsne(X = as.matrix(df_nums_adjscaled), perplexity = 40, max_iter = 5000, pca = FALSE, momentum = 0.75) -->
<!-- tsneFit -->
<!-- toPlot <- tbl_df(tsneFit$Y) -->
<!-- names(toPlot) <- c('X','Y') -->
<!-- xyplot(X~Y, data = toPlot) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_nums_adj %>% bind_cols(toPlot) %>%  -->
<!--   lattice::xyplot(Y~X,groups=q13_visitfreq_video,data=.,auto.key=list(columns=6)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_nums_adj %>% bind_cols(toPlot) %>% write_csv(path = '../data/tsne_out.csv', col_names = T) -->
<!-- ``` -->